Naturalness in U(1) extensions of the MSSM

### 2014-12-17 paper/e6ssm-tuning-paper.tex

\section{\label{sec:results}Results}
% Our results. Plots to go in here:
%    - tuning as a function of stop mass for the MSSM and E6SSM
Many recent papers have focussed on light stops as allowing natural
SUSY and there has been a lot of theoretical effort to find SUSY
models with a $125$ GeV Higgs boson while still having light stops
and experimental efforts to search for light stops.  This is entirely
appropriate, however the fine tning problem from the stops depends
crucially on the RG evolution from the GUT scale, since this is how
the squark masses feed into the soft Higgs masses that appear in the
EWSB conditions.

To illustrate this, in Fig. \ref{} we show the variation in fine tuning when we fix $\mu=-97$ GeV and $B=-85$ GeV, 
and scan over the stop masses and mixing, with $500 \textrm{ GeV} \leq m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and $-3810\textrm{ GeV}\leq A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV, $50$ TeV and $100$ TeV. 
However this tuning only really becomes large for very large stop masses.  On
the other hand if cuttoff is increased then stop masses of several
TeV can result in tuning of [insert] as shown in Fig.~\ref{}.\\ \\

@@

Therefore the tuning in these models depends on less assumptions than
the tuning which concerns people in the MSSM and the owest tuned
models are adderessed directly with $Z^\prime$ searches.

In the MSSM the tuning can also be attcked more directly by looking at
chargino mass limikts as this rasies $\mu$ directly.  [Stuff aboiut
  chargino limits stiuts, then show plots in MSSM and E6SSM]
%    - plots of chargino bound impacts

@@

As can be seen in Fig.~\ref{} the severity of the tunings varies quite a
bit.  This is because the charges appear as coefficients in front of
the $Z^\prime$ mass in the EWSB condition.  These cagres change the
value of the coefficient $d$ in Eq.~\ref{}.  The values of the
coefficient $d$ in each model is [insert], and this determines which
of the models are most tuned.\\ \\

% DH:: I didn't use the d notation in the section on EWSB, would this
% work to introduce it here instead?
While the $Z^\prime$ mass does set a lower limit on the tuning, the exact
value of this tuning depends quite strongly on the charges of the extra
$U(1)$ gauge symmetry it is associated with. This is illustrated in Fig~\ref{}, 
in which we show the fine tuning for several other commonly studied
$U(1)$ extensions for the same set of $Z^\prime$ masses as were considered in the
E$_6$SSM. As can be seen in Fig.~\ref{}, the severity of the tunings varies
quite a bit. This can be understood from the fact that the $D$-terms $D_{H_d}$, $D_{H_u}$
that enter the EWSB condition Eq. (\ref{eq:E6MZequation}) are proportional to
the $U(1)'$ charges. It is therefore possible to significantly reduce the size
of the additional $D$-term contributions to the EWSB conditions by choosing the $U(1)'$ 
charges appropriately. For $s\gg v$, as is
required in all of the $U(1)$ extended models considered in order to have $M_{Z^\prime}\geq 2.5$ TeV,
the $D$-term contribution to Eq. (\ref{eq:E6MZequation}) is well approximated by
\begin{equation*}
\frac{D_{H_d}-D_{H_u}\tan^2\beta}{\tan^2-1}\approx \frac{g_1'^2}{2}\tilde{Q}_Ss^2\left ( \frac{\tilde{Q}_1-\tilde{Q}_2\tan^2\beta}{\tan^2\beta-1}\right )\approx d(\theta,\tan\beta)\frac{M_{Z^\prime}^2}{2},\qquad d(\theta,\tan\beta)\equiv \left ( \frac{\tilde{Q}_1-\tilde{Q}_2\tan^2\beta}{\tilde{Q}_S\left ( \tan^2\beta - 1\right )}\right ) 
\end{equation*}
for models with $\tilde{Q}_S\neq 0$. Written in this form one might expect that the fine tuning
associated with this contribution will be reduced when this expression is 
minimised. This is indeed the case. For example, in the E$_6$SSM with $\theta=\arctan\sqrt{15}$ one finds that for $\tan\beta=10$, $d\approx 0.398$. This can be compared to the $U(1)_\psi$ ($\theta=\pi/2$) and $U(1)_\eta$ ($\theta=-\arctan\sqrt{5/3}$) models, for which $d=0.5$ and $d\approx 0.806$, respectively. The larger values of the $D$-term contributions in these models result in them having a larger fine tuning than in the E$_6$SSM, as is evident from Fig.~\ref{}. Of particular note is the inert $U(1)_I$ model ($\theta=\pi+\arctan\sqrt{3/5}$), which has $\tilde{Q}_2=0$, resulting in the coefficient $d$ being suppressed by $\tan^2\beta$. For this model we find that $d\approx-0.01$, and so the $D$-term contribution is much smaller than in the other models described. The fine tuning is as a result somewhat smaller, not being much greater than that found in the MSSM.

>>>>

\section{\label{sec:results}Results}
% Our results. Plots to go in here:
%    - tuning as a function of stop mass for the MSSM and E6SSM
Many recent papers have focussed on light stops as allowing natural
SUSY and there has been a lot of theoretical effort to find SUSY
models with a $125$ GeV Higgs boson while still having light stops
and experimental efforts to search for light stops.  This is entirely
appropriate, however the fine tuning problem from the stops depends
crucially on the RG evolution from the GUT scale, since this is how
the squark masses feed into the soft Higgs masses that appear in the
EWSB conditions.

To illustrate this, in Fig. \ref{} we show the variation in fine
tuning when we fix $\mu=-97$ GeV and $B=-85$ GeV, and scan over the
stop masses and mixing, with $500 \textrm{ GeV} \leq
m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and $-3810\textrm{ GeV}\leq
A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV, $50$ TeV and $100$ TeV.
However this tuning only really becomes large for very large stop
masses.  On the other hand if the cuttoff is increased then stop masses of
several TeV can result in tuning of [insert] as shown in
Fig.~\ref{}.\\ \\

@@

Therefore the tuning in these models depends on less assumptions than
the tuning which concerns people in the MSSM and the lowest tuned
models are adderessed directly with $Z^\prime$ searches.

In the MSSM the tuning can also be attcked more directly by looking at
chargino mass limikts as this rasies $\mu$ directly.  [Stuff about
  chargino limits, then show plots in MSSM and E6SSM]
%    - plots of chargino bound impacts

@@

As can be seen in Fig.~\ref{} the severity of the tunings varies quite a
bit.  This is because the charges appear as coefficients in front of
the $Z^\prime$ mass in the EWSB condition.  These charges change the
value of the coefficient $d$ in Eq.~\ref{}.  The values of the
coefficient $d$ in each model is $\{-0.01, 0.40, 0.50, 0.81 \}$ for
$\{U(1)_I, U(1)_N, U(1)_\psi, U(1)_\eta \}$ and this determines which
of the models are most tuned.

% DH:: I didn't use the d notation in the section on EWSB, would this
% work to introduce it here instead?
While the $Z^\prime$ mass does set a lower limit on the tuning, the exact
value of this tuning depends quite strongly on the charges of the extra
$U(1)$ gauge symmetry it is associated with. This is illustrated in Fig~\ref{}, 
in which we show the fine tuning for several other commonly studied
$U(1)$ extensions for the same set of $Z^\prime$ masses as were considered in the
E$_6$SSM. As can be seen in Fig.~\ref{}, the severity of the tunings varies
quite a bit. This can be understood from the fact that the $D$-terms $D_{H_d}$, $D_{H_u}$
that enter the EWSB condition Eq. (\ref{eq:E6MZequation}) are proportional to
the $U(1)'$ charges. It is therefore possible to significantly reduce the size
of the additional $D$-term contributions to the EWSB conditions by choosing the $U(1)'$ 
charges appropriately. For $s\gg v$, as is
required in all of the $U(1)$ extended models considered in order to have $M_{Z^\prime}\geq 2.5$ TeV,
the $D$-term contribution to Eq. (\ref{eq:E6MZequation}) is well approximated by
\begin{equation*}
\frac{D_{H_d}-D_{H_u}\tan^2\beta}{\tan^2-1}\approx \frac{g_1'^2}{2}\tilde{Q}_Ss^2\left ( \frac{\tilde{Q}_1-\tilde{Q}_2\tan^2\beta}{\tan^2\beta-1}\right )\approx d(\theta,\tan\beta)\frac{M_{Z^\prime}^2}{2},\qquad d(\theta,\tan\beta)\equiv \left ( \frac{\tilde{Q}_1-\tilde{Q}_2\tan^2\beta}{\tilde{Q}_S\left ( \tan^2\beta - 1\right )}\right ) 
\end{equation*}
for models with $\tilde{Q}_S\neq 0$. Written in this form one might expect that the fine tuning
associated with this contribution will be reduced when this expression is 
minimised. This is indeed the case. For example, in the E$_6$SSM with $\theta=\arctan\sqrt{15}$ one finds that for $\tan\beta=10$, $d\approx 0.398$. This can be compared to the $U(1)_\psi$ ($\theta=\pi/2$) and $U(1)_\eta$ ($\theta=-\arctan\sqrt{5/3}$) models, for which $d=0.5$ and $d\approx 0.806$, respectively. The larger values of the $D$-term contributions in these models result in them having a larger fine tuning than in the E$_6$SSM, as is evident from Fig.~\ref{}. Of particular note is the inert $U(1)_I$ model ($\theta=\pi+\arctan\sqrt{3/5}$), which has $\tilde{Q}_2=0$, resulting in the coefficient $d$ being suppressed by $\tan^2\beta$. For this model we find that $d\approx-0.01$, and so the $D$-term contribution is much smaller than in the other models described. The fine tuning is as a result somewhat smaller, not being much greater than that found in the MSSM. 
>>>>>>> 9210c662723efc5e3c4599b1702311631b8f859c
